Compositional Quantitative Resource Analysis of C Programs

### 2014-06-24 paper/main.tex
small fixes to the call rule
new inference rules

\begin{figure*}[t]
\begin{mathpar}
%
\Rule{Q:Skip}
{}
{ \Gamma; B; R; Q \vdash \code{skip} \dashv Q; \Gamma }
%
\and \Rule{Q:Break}
{}
{ \Gamma; B; R; B + M_b \vdash \code{break} \dashv Q'; \Gamma' }
%
\and \Rule{Q:Return}
{}
{ \Gamma; B; R; R \, x + M_r \vdash \code{return}~x \dashv Q'; \Gamma' }
%
\and \Rule{Q:Assert}
{}
{ \Gamma; B; R; Q + M_a \vdash \code{assert}~e \dashv Q; \Gamma, e }
%
\and \Rule{Q:Loop}
{ \Gamma; Q'; R; Q \vdash S \dashv M_l + Q; \Gamma }
{ \Gamma; B; R; Q \vdash \code{loop}~S \dashv Q'; \code{inv}_S~\Gamma }
%
\and \Rule{Q:Set}
{ q'_{xy}, q'_{yx} \in \mathbb Q_0^+
\\ \forall u. (q_{yu} = q'_{xu} + q'_{yu} \land q_{uy} = q'_{ux} + q'_{uy})
}
{ \Gamma[x/y]; B; R; Q + M_s \vdash x \gets y \dashv Q'; \Gamma }
%
\and \Rule{Q:IncrPos}
{ \Gamma \models 0 \le y
\\ \mathcal U = \left\{ u \mid \Gamma \models x + y \in \inter x u \right\}
\\\\ \textstyle q'_{0y} = q_{0y}
      + \sum_{u \in \mathcal U} q_{xu}
      - \sum_{v \not\in \mathcal U} q_{vx}
}
{ \textstyle
  \Gamma[x/x + y]; B; R; Q + M_s \vdash x \gets x + y \dashv Q'; \Gamma
}
%
\and \Rule{Q:IncrNeg}
{ \Gamma \models 0 \le y
\\ \mathcal U = \left\{ u \mid \Gamma \models x - y \in \inter u x \right\}
\\\\ \textstyle q'_{y0} = q_{y0}
      + \sum_{u \in \mathcal U} q_{ux}
      - \sum_{v \not\in \mathcal U} q_{xv}
}
{ \textstyle
  \Gamma[x/x - y]; B; R; Q + M_s \vdash x \gets x - y \dashv Q'; \Gamma
}
%
\and \Rule{Q:Incr}
{ \textstyle q'_{0y} = q_{0y} - \sum_{v} q_{vx}
\\ \textstyle q'_{y0} = q_{y0} - \sum_{v} q_{xv}
}
{ \textstyle
  \Gamma[x/x \pm y]; B; R; Q + M_s \vdash x \gets x \pm y \dashv Q'; \Gamma
}
%
\and \Rule{Q:If}
{ \Gamma, e; B; R; Q - M_c^1 \vdash S_1 \dashv Q'; \Gamma'
\\ \Gamma, \neg e; B; R; Q - M_c^2 \vdash S_2 \dashv Q'; \Gamma'
}
{ \Gamma; B; R; Q \vdash \code{if}(e)~S_1~\code{else}~S_2 \dashv Q'; \Gamma' }
%
\and\Rule{Q:Seq}
{  \Gamma; B; R; Q - M_s^1 \vdash S_1 \dashv Q'; \Gamma'
\\ \Gamma'; B; R; Q' - M_s^2 \vdash S_2 \dashv Q''; \Gamma''
}
{ \Gamma; B; R; Q \vdash S_1;S_2 \dashv Q''; \Gamma'' }
%
\and\Rule{Q:Call}
{ (\vec y, s, \Gamma_f, Q_f, Q_f', \Gamma_f') \in \Delta(f)
\\ \text{Loc} = \text{Locals}(Q)
\\ \forall i \neq j . \, x_i \neq x_j
\\ c \in \mathbb Q_0^+
\\\\ Q \ge P + S
\\ P' + S \ge Q'
\\ R = Q_f[\vec y / \vec x]
\\ R' = Q'_f[s/r]
\\ \forall i \in \dom{R} . \, p_i = r_i
\\ \forall i \in \dom{R'} .\, p'_i = r'_i 
\\ \forall i \not\in \dom{R'} .\, p'_i = 0
\\ \forall i \not\in \text{Loc}
  .\, s_i = 0
}
{ \Gamma_f[\vec y / \vec x] \land \Gamma_{\text{Loc}}; B; R; Q + c
  \vdash r \gets f(\vec x) \dashv
  Q' + c; \Gamma_f'[s/r] \land \Gamma_{\text{Loc}}
}
%
\and \Rule{Q:Weak}
{ \Gamma_1 \models \Gamma_2
\\ Q_1 \succeq_\Gamma Q_2
\\ \Gamma_2; B; R; Q_2 \vdash S \dashv Q'_2; \Gamma'_2
\\ Q'_2 \succeq_\Gamma Q'_1
\\ \Gamma'_2 \models \Gamma'_1
}
{ \Gamma_1; B; R; Q_1 \vdash S \dashv Q'_1; \Gamma'_1 }
%
\and \Rule{Relax}
{ \mathcal L = \{ xy \mid \exists l_{xy},~\Gamma \models l_{xy} \le |\inter x y| \}
\and \mathcal U = \{ xy \mid \exists u_{xy},~\Gamma \models |\inter x y| \le u_{xy} \}
\\ \forall i \in \mathcal U.\, q'_i \ge q_i - p_i
\\ \forall i \in \mathcal L.\, q'_i \ge q_i + p_i
\\ \forall i \not\in \mathcal U \cup \mathcal L.\, q'_i \ge q_i
}
{ Q' + \sum_{i \in \mathcal U} u_i p_i - \sum_{i \in \mathcal L} l_i p_i  \succeq_\Gamma Q }
%
\end{mathpar}
\caption{Abstract proof rules of the quantitative analysis}
\label{fig:proof}
\end{figure*}

>>>>

\begin{figure*}[t]
\begin{mathpar}
%
\Rule{Q:Skip}
{}
{ B; R; (\Gamma, Q) \vdash \code{skip} \dashv (\Gamma, Q) }
%
\and \Rule{Q:Break}
{}
{  (\Gamma, Q_B); R; (\Gamma, Q_B {+} M_b) \vdash \code{break} \dashv (\Gamma', Q') }
%
\and \Rule{Q:Return}
{ P = Q_R[\Vret/x]  
\\ \forall i \in \dom{P} . \, p_i = q_i
}
{ B; (\Gamma,Q_R); (\Gamma, Q{+} M_r) \vdash \code{return}~x \dashv (\Gamma',Q') }
%
\and \Rule{Q:Assert}
{}
{ B; R; (\Gamma, Q {+} M_a) \vdash \code{assert}~e \dashv (\Gamma \land e, Q) }
%
\and \Rule{Q:Loop}
{ (\Gamma', Q'); R; (\Gamma,Q) \vdash S \dashv (\Gamma,  Q {+} M_l) }
{ B; R; (\Gamma, Q) \vdash \code{loop}~S \dashv (\Gamma', Q') }
%
\and \Rule{Q:Set}
{ q'_{xy}, q'_{yx} \in \mathbb Q_0^+
\\ \forall u. (q_{yu} = q'_{xu} + q'_{yu} \land q_{uy} = q'_{ux} + q'_{uy})
}
{ B; R; (\Gamma[x/y], Q {+} M_s) \vdash x \gets y \dashv (\Gamma, Q') }
%
\and \Rule{Q:IncrPos}
{ \Gamma \models 0 \le y
\\ \mathcal U = \left\{ u \mid \Gamma \models x + y \in \inter x u \right\}
\\\\ \textstyle q'_{0y} = q_{0y}
      + \sum_{u \in \mathcal U} q_{xu}
      - \sum_{v \not\in \mathcal U} q_{vx}
}
{ \textstyle
  B; R; (\Gamma[x/x {+} y], Q {+} M_s) \vdash x \gets x + y \dashv (\Gamma, Q')
}
%
\and \Rule[rightskip=1cm]{Q:IncrNeg}
{ \Gamma \models 0 \le y
\\ \mathcal U = \left\{ u \mid \Gamma \models x - y \in \inter u x \right\}
\\\\ \textstyle q'_{y0} = q_{y0}
      + \sum_{u \in \mathcal U} q_{ux}
      - \sum_{v \not\in \mathcal U} q_{xv}
}
{ \textstyle
  B; R; (\Gamma[x/x {-} y],Q {+} M_s) \vdash x \gets x - y \dashv (\Gamma, Q')
}
%
\and \Rule{Q:Incr}
{ \textstyle q'_{0y} = q_{0y} - \sum_{v} q_{vx}
\\ \textstyle q'_{y0} = q_{y0} - \sum_{v} q_{xv}
}
{ \textstyle
  B; R; (\Gamma[x/x \pm y], Q {+} M_s) \vdash x \gets x \pm y \dashv (\Gamma, Q')
}
%
\and \Rule{Q:If}
{ B; R; (\Gamma \land e, Q {-} M_c^1) \vdash S_1 \dashv (\Gamma',Q')
\\ B; R; (\Gamma \land \neg e, Q {-} M_c^2) \vdash S_2 \dashv (\Gamma',Q')
}
{ B; R; (\Gamma, Q) \vdash \code{if}(e)~S_1~\code{else}~S_2 \dashv (\Gamma',Q') }
%
\and\Rule{Q:Seq}
{  B; R; (\Gamma, Q {-} M_s^1) \vdash S_1 \dashv (\Gamma', Q')
\\ B; R; (\Gamma', Q' {-} M_s^2) \vdash S_2 \dashv (\Gamma'',Q'')
}
{ B; R; (\Gamma, Q) \vdash S_1;S_2 \dashv (\Gamma'',Q'') }
%
\and\Rule{Q:Call}
{ (\Gamma_f, Q_f,  \Gamma_f', Q_f') \in \Delta(f)
\and \text{Loc} = \text{Locals}(Q)
\and \forall i \neq j . \, x_i \neq x_j
\and c \in \mathbb Q_0^+
\\ Q = P + S
\and Q' = P' + S
\and U = Q_f[\Vargs / \vec x]
\and U' = Q'_f[\Vret/r]
\\ \forall i \in \dom{U} . \, p_i = u_i
\and \forall i \in \dom{U'} .\, p'_i = u'_i 
\and \forall i \not\in \dom{U'} .\, p'_i = 0
\and \forall i \not\in \text{Loc}
  .\, s_i = 0
}
{ B; R; (\Gamma_f[\Vargs / \vec x] \land \Gamma_{\text{Loc}}, Q {+} c)
  \vdash r \gets f(\vec x) \dashv
  (\Gamma_f'[\Vret/r] \land \Gamma_{\text{Loc}}, Q' {+} c)
}
%
\and \Rule{Q:Weak}
{ \Gamma_1 \models \Gamma_2
\\ Q_1 \succeq_{\Gamma_1} Q_2
\\ B; R; (\Gamma_2,Q_2) \vdash S \dashv (\Gamma'_2, Q'_2)
\\ Q'_2 \succeq_{\Gamma_2'} Q'_1
\\ \Gamma'_2 \models \Gamma'_1
}
{ B; R; (\Gamma_1,Q_1) \vdash S \dashv (\Gamma'_1, Q'_1) }
%
\and \Rule{Relax}
{ \mathcal L = \{ xy \mid \exists l_{xy},~\Gamma \models l_{xy} \le |\inter x y| \}
\and \mathcal U = \{ xy \mid \exists u_{xy},~\Gamma \models |\inter x y| \le u_{xy} \}
\\ \forall i \in \mathcal U.\, q'_i \ge q_i - p_i
\\ \forall i \in \mathcal L.\, q'_i \ge q_i + p_i
\\ \forall i \not\in \mathcal U \cup \mathcal L.\, q'_i \ge q_i
}
{ Q' + \sum_{i \in \mathcal U} u_i p_i - \sum_{i \in \mathcal L} l_i p_i  \succeq_\Gamma Q }
%
\and \Rule{Q:Abst}
{ \fenv f = (\vec y, S_f)
\\ B; (\Gamma_f', Q'_f); (\Gamma_f[\Vargs/\vec y], Q_f[\Vargs/ \vec y] ) \vdash S \dashv (\Gamma', Q')
}
{ (\Gamma_f, Q_f, \Gamma_f', Q_f') \in \Delta(f) }
%
\end{mathpar}
\caption{Inference rules of the quantitative analysis}
\label{fig:proof}
\end{figure*}
